### 2014-03-06 paper/paper.tex
convert paper to uai proceeding

% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two
% column, 8 1/2 by 11 inch proceedings format.

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{latexsym,amsbsy,amssymb,color,xspace, booktabs}

\include{macros}
\include{local_macros}

\title{Stochastic Multiple Output Gaussian Processes}

\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

\begin{document}

\maketitle

\begin{abstract}
The Abstract paragraph 
\end{abstract}

\section{INTRODUCTION}
\input{intro}

\subsubsection{Citations in Text}

Citations within the text should include the author's last name and
year, e.g., (Cheesman, 1985). Reference style should follow the style
that you are used to using, as long as the citation style is
consistent.

\section{MODEL DESCRIPTION}
\input{model}

\subsection{THIS SHOULD BE CAP}

\section{INFERENCE}
\input{inference}

\section{EXPERIMENTS}
\input{experiments}

\section{DISCUSSION}
\input{discussion}

\subsubsection{Figures}

All artwork must be centered, neat, clean, and legible. Figure number
and caption always appear below the figure.  Leave 2 line spaces
between the figure and the caption. The figure caption is initial caps
and each figure numbered consecutively.

Make sure that the figure caption does not get separated from the
figure. Leave extra white space at the bottom of the page rather than
splitting the figure and figure caption.
\begin{figure}[h]
\vspace{1in}
\caption{Sample Figure Caption}
\end{figure}

\subsubsection{Tables}

\begin{table}[h]
\caption{Sample Table Title}
\label{sample-table}
\begin{center}
\begin{tabular}{ll}
\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION} \\
\hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\newpage
\section{APPENDIX}
Perhaps this should go to a supplementary material.
\input{appendix}

%\subsubsection*{Acknowledgements}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}