### 2014-03-06 paper/paper.tex
paper

% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two
% column, 8 1/2 by 11 inch proceedings format.

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{latexsym,amsbsy,amssymb,color,xspace, booktabs}

\include{macros}
\include{local_macros}

\title{Stochastic Multiple Output Gaussian Processes}

\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

\begin{document}

\maketitle

\begin{abstract}
The Abstract paragraph 
\end{abstract}

\section{INTRODUCTION}
\input{intro}

\subsubsection{Citations in Text}

Citations within the text should include the author's last name and
year, e.g., (Cheesman, 1985). Reference style should follow the style
that you are used to using, as long as the citation style is
consistent.

\section{MODEL DESCRIPTION}
\input{model}

\subsection{THIS SHOULD BE CAP}

\section{INFERENCE}
\input{inference}

\section{EXPERIMENTS}
\input{experiments}

\section{DISCUSSION}
\input{discussion}
\newpage
\section{APPENDIX}
Perhaps this should go to a supplementary material.
\input{appendix}

%\subsubsection*{Acknowledgements}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}