### 2014-03-06 paper/note.tex

\documentclass{article} % For LaTeX2e
\usepackage[a4paper,margin=1.5in]{geometry}
\usepackage[mathlines]{lineno}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{latexsym,amsbsy,amssymb,color,xspace, booktabs}

\include{macros}
\include{local_macros}
\def\linenumberfont{\normalfont\small\sffamily}

\title{Multiple Output Gaussian Processes Regression}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle
\linenumbers

\section{Model Description}
We describe a multiple-output regression model.
Suppose that we have inputs $\X = \{\x_n\}_{n=1}^N$ and outputs $\y = \{\y_n\}_{n=1}^N$, with $\x_n \in \calR^D$ and $\y_n \in \calR^P$, i.e. the output is $P$-dimensional.
For ease of exposition, we assume that there are no missing values in any output dimension.
The most dominant approach in GP-based multiple output regression is to compose the different outputs as some \textit{linear} mixing (combination) of some basic processes (random functions).
Our model is also underpinned by this composition approach.

\noindent Consider the semiparametric latent factor models (SLFM) \cite{teh-et-al-aistats-05}.
It assumes $Q$ basic independent processes, $g_j(\x) \sim \GP(0, k(\x,\x';\vectheta_j))$, where $\vectheta_j$ is the hyperparameters of the covariance function of $g_j(\x)$.
Each (latent) output is a weighted combination of these basic processes, i.e.
\begin{align}
f_i(\x) = \sum_{j=1}^Q w_{ij} g_j(\x), \quad i = 1 \hdots P,
\end{align}
In \cite{seeger2005semiparametric}, an extra degree of freedom is given by allowing each output $i$ to have its own process $h_i(\x) \sim \GP(0, k(\x,\x'; \vectheta^h_i)$ leading to
\begin{align}
f_i(\x) = \sum_{j=1}^Q w_{ij} g_j(\x) + h_i(\x), 
\end{align}
We emphasize that, although $k$ denotes covariance functions in general, each independent process can have its own covariance function.

\noindent 
\textbf{Augmented sparse GPs}
Toward scalable modeling, we replace standard GPs with sparse GPs augmented with \textit{different} set of inducing inputs.
This adds much flexibility to the model as each of the shared process $g_j(\x)$ can model a different pattern in the data with its own covariance function and inducing inputs. The roles of $g_j(\x)$ and $h_i(\x)$ can be quite different, so it is necessary that each has its own inducing inputs.
Furthermore, $g_j(\x)$ can be seen as a \textit{global} function operating on the entire input space (of all output dimensions), while each $h_i(\x)$ operates only on the inputs of the $i$-th output, which can be a subspace of the input.

%However, to really address the issue of scalability, this model is built upon sparse GPs from the ground up.

\subsection{Prior and Likelihood}
In this section we specify the prior and likelihood of this model. 
\newcommand{\Zj}{\Z_j}
\newcommand{\Zhi}{\Z^h_i}
Let the set of inducing inputs for $g_j(\x)$ and $h_i(\x)$ be $\Z_j$ and $\Z^h_i$, and the corresponding inducing points $\u_j$ and $\v_i$, respectively.
We denote the collective variables: $\g = \{\g_j\}$, $\h = \{\h_i \}$, $\u = \{\u_j\}$, $\v = \{\v_i\}$, $\Z = \{\Zj\}$, and $\Z^h = \{\Zhi \}$ where $\g_j = \{g_j(\x_n)\}$, $\h_i = \{h_i(\x_n)\}$. 
The subscripts are $i = 1 \hdots P$, $j = 1 \hdots Q$, and $n = 1 \hdots N$.

\noindent
\textbf{Prior}
The prior of the augmented model is given by:
\begin{align}
p(\g | \u) &= \prod_{j=1}^Q p(\g_j | \u_j) = \prod_{j=1}^Q \Normal(\g_j; \BigMu_j, \tilde{\K}_j )\\
p(\u) &= \prod_{j=1}^Q p(\u_j) = \prod_{j=1}^Q \Normal(\u_j; \vec{0}, k(\Zj, \Zj)) \\
p(\h | \v) &= \prod_{i=1}^P p(\h_i | \v_i) = \prod_{i=1}^P \Normal(\h_i; \BigMu^h_i, \tilde{\K}^h_i)\\
p(\v) &= \prod_{i=1}^P p(\v_i) = \prod_{i=1}^P \Normal(\v_i; \vec{0}, k(\Zhi, \Zhi)),
\end{align}
where $\BigMu_j = k(\X,\Zj)k(\Zj,\Zj)^{-1}\u$ and $\tilde{\K}_j = k_j(\X,\X) - k(\X,\Zj)k(\Zj,\Zj)^{-1}k(\Zj,\X)$ and $\BigMu^h_i, \tilde{\K}^h_i$ are similarly defined.
To avoid notational clutter, we omit the subscripts $j,h,i$ from the kernels $k_j(\cdot,\cdot)$ and $k^h_i(\cdot,\cdot)$ when it is clear from the parameters inside the parentheses which covariance function is in action.

\noindent
\textbf{Likelihood}
The likelihood as usual follows the standard iid Gaussian likelihood,
\begin{align}
p(\y | \g, \h ) = \prod_{i=1}^P p( \y_i ; \g, \h_i) = \prod_{i=1}^P \Normal( \y_i ; \sum_{j=1}^Q w_{ij} \g_j + \h_i, \beta_i^{-1} \I).
\end{align}

\subsection{Variational Inference in Sparse GPs Revisited}
In this section we review inference in sparse GPs (as presented in Titsias).
The posterior in an augmented model is $p(\f, \u | \y) = p(\f | \u, \y) p(\u | \y)$.
The key property of such augmented GPs  is the notation of \textit{sufficient statistics}: given the inducing points $\u$, the latent values $\f$ are independent with any other set of latent values (e.g. the test set).
In the optimal setting when $\u$ is the sufficient statistics of $\f$, it should hold that $p(\f | \u, \y) = p(\f | \u)$ as $\y$ is only the noisy version of $\f$.
This leads to choosing a variational approximation of the posterior which factorizes as $q(\f, \u | \y) = p(\f | \u) q(\u | \y)$.
Since the conditional $p(\f | \u)$ is known, variational inference becomes learning an optimal posterior $q(\u | \y)$ only. \\

\noindent Also as a consequence of sufficient statistics, the approximate prediction for test targets $\vfstar$ at test inputs $\X_*$ is
\begin{align}
\nonumber
p(\vfstar | \y, \X_*) &= \int p(\vfstar | \f, \u, \X_*) q(\f, \u | \y) \der \f \der \u \\
\nonumber
&= \int p(\vfstar | \u) q(\u| \y) p(\f | \u) \der \f \der \u \\ \nonumber
&= \int p(\vfstar | \u) q(\u| \y) \der \u \\
\label{eq:sorprediction}
&= \Normal(\vfstar; \bs{\mu_*},\vec{s_*})
\end{align}
where,
\begin{align}
\nonumber
\bs{\mu_*} &= \K_{*z} \K_{zz}^{-1}\m \\ 
\nonumber
\S_* &= \K_{**} - \K_{*z} \left(\K_{zz}^{-1} - \K_{zz}^{-1} \S \K_{zz}^{-1} \right) \K_{*z}^T.
\end{align}
 Here $\K_{*z}$ is the covariance matrix between test and inducing inputs, $\K_{zz}$ is the covariance matrix of the inducing inputs.

\subsection{Variational Inference}
\newcommand{\ug}{\u_g}
\newcommand{\uh}{\u^h}
\newcommand{\mgj}{\m_j}
\newcommand{\mhi}{\m^h_i}
\newcommand{\Sgj}{\S_j}
\newcommand{\Shi}{\S^h_i}
Our goal of inference is to find the posterior $p(\g, \h, \u, \v | \y)$. 
Following the previous discussion, we assume a variational distribution which factorizes as:
\begin{align}
\nonumber
q(\g, \h, \u, \v | \y)
\nonumber
 &= p(\g|\u) p(\h|\v) q(\u,\v)  \\
 &= p(\g|\u) p(\h|\v) \prod_{j=1}^Q q(\u_j) \prod_{i=1}^P  q(\v_i)
\end{align}
Since the conditionals $p(\g | \u)$ and $p(\h | \v)$ are given, we need only to find the optimum $q(\u_j)$ and $q(\v_i)$.
Let $q(\u_j) = \Normal(\u_j; \mgj, \Sgj)$ and $q(\v_i) = \Normal(\v_i; \mhi, \Shi)$. \\

\noindent
To find the optimum $q(\u, \v)$ we optimize the evidence lower bound (ELBO) of the log marginal,
\begin{align}
\nonumber
\log p(\y) \ge& \int q(\u, \v) \log \frac{p(\y | \u, \v) p(\u, \v)}{q(\u, \v)} \der \u \der \v \\
\nonumber
=& \int q(\u, \v) \log p(\y | \u, \v)  \der \u \der \v 
+ \int q(\u, \v) \log \frac{p(\u, \v)}{q(\u, \v)} \der \u \der \v \\
\label{eq:elbo}
=& \int q(\u, \v) \log p(\y | \u, \v)  \der \u \der \v 
- \left(\sum_{j=1}^Q \KL[q(\u_j) || p(\u_j)] + \sum_{i=1}^P \KL[q(\u_i) || p(\u_i)] \right),
\end{align}
where the last equality occurs because both of $q(\u, \v)$ and $p(\u, \v)$ fully factorize.
Since $q(\u_j), q(\v_i), p(\u_j), p(\v_i)$ are all multivariate Gaussian distributions, the KL divergences are analytically tractable and require $\calO(M^3)$ computation, where $M$ is the largest number of inducing inputs (recall that $g_j(\x)$ and $h_i(\x)$ use separate set of inducing inputs). \\

\noindent To compute the ELBO we first focus on the term:
\newcommand{\llangle}{\left\langle}
\newcommand{\rrangle}{\right\rangle}
\begin{align}
\nonumber
\log p(\y | \u, \v)
 &= \log \Eb{p(\y | \g, \h)}_{p(\g,\h | \u, \v)} \\
 \nonumber
&\ge \Eb{\log p(\y | \g, \h)}_{p(\g,\h | \u, \v)} \quad &\text{(Jensen's inequality)} \\
&= \sum_{i=1}^P \sum_{n=1}^N \Eb{\log p(y_{in} | \g_n, h_{in}) }_{p(\g_n | \u) p(\h_{in} | \v_i)}, \quad 
 &\text{(factorized likelihood)}
\end{align}
where $\g_n = \{g_{jn} = (\g_j)_n\}_{j=1}^Q$.
Each individual term $l_{in} \define \Eb{\log p(y_{in} | \g_n, h_{in}) }_{p(\g_n | \u) p(\h_in | \v_i)}$ can be computed using the identity in eq. \ref{eq:identity} given in the appendix to give:
\begin{align}
\nonumber
l_{in} &= \int \log p(y_{in} | \g_n, h_{in}) \prod_{j=1}^Q p(g_{jn} | \u_j) p(h_{in} | \u_i) \der \g_n \der h_{in} \\
&= \log \Normal(y_{in}; \sum_{j=1}^Q w_{ij} \mu_{jn}+ \mu^h_{in}, \beta_i^{-1})
- \frac{1}{2} \beta_i \sum_{j=1}^Q w_{ij}^2 \tilde{k}_{nn} 
- \frac{1}{2} \beta_i \tilde{k}^h_{inn},
\end{align}
where $\tilde{k}_{nn} = (\tilde{\K})_{nn}, \tilde{k}^h_{inn} = (\tilde{\K}^h_i)_{nn}, \mu_{jn} = (\Mu_j)_n, \text{ and } \mu^h_{in} = (\Mu^h_i)_n$.

% previous detailed derivation
%\begin{align}
%\nonumber
%l_{in} &= \int \log p(y_{in} | g_n, h_{in}) p(\g | \ug) p(\h_i | \u_i) \der \g \der \h_i \\
%\nonumber
%&= \int \log \Normal(y_{in} ; w_i g_n + h_{in}, \beta_i^{-1}) 
%\Normal(g_n ; \mu_{gn}, \tilde{k}_{gnn})
%\Normal(h_{in} ; \mu_{in}, (\tilde{k}_{inn}) \der g_n \der h_{in} \\
%\nonumber
%&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} \int (y_{in} - w_i g_n - h_{in}) \beta_i (y_{in} - w_i g_n - h_{in})
%\Normal(g_n ; \mu_{gn}, \tilde{k}_{gnn})
%\Normal(h_{in} ; \mu_{in}, \tilde{k}_{inn}) \der g_n \der h_{in} \\
%\nonumber
%&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} \int \left[(y_{in} - h_{in} - w_i \mu_{gn}) \beta_i (y_{in} - h_{in} - w_i \mu_{gn}) + w_i^2 \beta_i \tilde{k}_{gnn} \right] 
%\Normal(h_{in} ; \mu_{in}, \tilde{k}_{inn}) \der h_{in} \\
%\nonumber
%&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} w_i^2 \beta_i \tilde{k}_{gnn}
%- \frac{1}{2} \beta_i \tilde{k}_{inn} - \frac{1}{2} (y_{in} - w_i \mu_{gn} - \mu_{in}) \beta_i (y_{in} - w_i \mu_{gn} - \mu_{in}) \\
%&= \log \Normal(y_{in}; w_i \mu_{gn} + \mu_{in}, \beta_i^{-1})  - \frac{1}{2} \beta_i (w_i^2 \tilde{k}_{gnn} + \tilde{k}_{inn}),
%\end{align}

\noindent Substituting $l_{in}$ into equation \ref{eq:elbo} and carrying out the integral using the identity in eq. \ref{eq:identity} we get:
\begin{align}
\nonumber
\log p(\y)
\ge& \sum_{i=1}^P \sum_{n=1}^N
\bigg( \log \Normal(y_{in}; \tilde{\mu}_{in}, \beta_i^{-1})
          - \frac{1}{2} \beta_i \sum_{j=1}^Q w_{ij}^2        \tilde{k}_{nn} - \frac{1}{2} \beta_i \tilde{k}^h_{inn}
 \\ \nonumber
         &\quad \quad \quad \quad - \frac{1}{2} \beta_i \trace  \sum_{j=1}^Q w_{ij}^2 \S_j \mat{\Lambda}_{jn} - \beta_i \frac{1}{2} \trace \S^h_i \mat{\Lambda}_{in} 
\bigg) \\
&- \left(\sum_{j=1}^Q \KL[q(\u_j) || p(\u_j)] + \sum_{i=1}^P \KL[q(\v_i) || p(\v_i)] \right) \define \calL,
\end{align}
where 
%TODO \Lambda_gn depends on the output i so need a better notation
\newcommand{\Zg}{\Z_g}
\newcommand{\Zi}{\Z_i}
\begin{align}
\tilde{\mu}_{in}
&= \sum_{j=1}^Q w_{ij} k(\x_n, \Zj)k(\Zj,\Zj)^{-1}\m_j + k(\x_n, \Zhi)k(\Zhi,\Zhi)^{-1}\mhi \\
\mat{\Lambda}_{jn}
&= k(\Zj,\Zj)^{-1} k(\Zj, \x_n) k(\x_n, \Zj) k(\Zj,\Zj)^{-1} \\
\mat{\Lambda}_{in}
&= k(\Zhi,\Zhi)^{-1} k(\Zhi, \x_n) k(\x_n, \Zhi) k(\Zhi,\Zhi)^{-1}.
\end{align}

\noindent Notice that this ELBO clearly generalizes the standard GP regression. In particular, setting $P = Q = 1$, $w_i = 1$ and $h_i(\x) = 0$ we recover the bound in Hensman et al \cite{hensmangaussian}.
Due to the decomposition of this bound, we can use stochastic gradient descent to learn the variational parameters.

\subsubsection{Variational Parameters Derivatives}
\newcommand{\oi}{\vec{o}_i}
% some notation
Before diving into the details, we first define the indexing operator of a matrix: $\B(\vec{r},\vec{c})$ extracts the submatrix in rows $\vec{r}$ and columns $\vec{c}$ of $\B$.
To index all columns we use $\B(\vec{r},:)$ and similarly for all rows $\B(:,\vec{c})$.
Readers familiar with this operator will recognize that this is the MATLAB indexing operator.

The optimal posteriors are found by setting the gradients of the lowerbound $\calL$ wrt to the parameters of $q(\u_j)$ and $q(\v_i)$.
Recall that in the model, different outputs can be observed at different inputs (i.e. the case of missing values).
Let $\oi$ be the indice of the observed inputs of the output dimension $i$.
We denote its set of observed inputs and targets as: $\X_i = \X(\oi,:)$ and $\y_i = y_{i}(\oi)$.

\noindent As a function of the parameters of $q(\u_j)$, the lowerbound $\calL$ is:
\newcommand{\Ahi}{\A^h_i}
\begin{align}
\nonumber
\calL^g_j \define&
 \sum_{i=1}^P \log \Normal(\y_i; \sum_{j=1}^Q w_{ij} \A_j(\oi,:) \m_j + \Ahi \mhi, \beta_i^{-1} \I)  \\
 \nonumber
 &- \frac{1}{2} \sum_{i=1}^P \bigg(\beta_i \trace w_{ij}^2 \tilde{\K}_j(\oi,\oi) 
 + \beta_i \trace w_{ij}^2 \S_j \A_j(\oi,:)^T \A_j(\oi,:) \bigg)
 \\
 &- \frac{1}{2} \log |k(\Zj,\Zj) \S_j^{-1}| -\frac{1}{2} \trace k(\Zj,\Zj)^{-1} (\m_j \m_j^T + \S_j) ,
\end{align}
where $\A_j = k(\X,\Zj)k(\Zj,\Zj)^{-1}$, which gives $\A_j(\oi,:) = k(\X_i,\Zj) k(\Zj,\Zj)^{-1}$, and  
$\Ahi = k(\X_i,\Zhi)k(\Zhi,\Zhi)^{-1}$. \\

%------------------------------------------
% derivatives of q(u_j)
\newcommand{\Lgj}{\calL^g_j}
\newcommand{\ynoj}{\y_i^{\backslash j}}
\noindent The derivatives of $\Lgj$ wrt $\m_j$ and $\S_j$ are given by:
\begin{align}
\deriv{\Lgj}{\m_j}
& = \sum_{i=1}^P \beta_i w_{ij} \A_j(\oi,:)^T \ynoj - \bigg[k_g(\Zg,\Zg)^{-1} + \sum_{i=1}^P \beta_i w_{ij}^2 \A_j(\oi,:)^T \A_j(\oi,:) \bigg] \m_j \\
\deriv{\Lgj}{\S_j} 
&= \frac{1}{2} \S_j^{-1} - \frac{1}{2} \bigg[ k(\Zj,\Zj)^{-1} + \sum_{i=1}^P \beta_i w_{ij}^2 \A_j(\oi,:)^T \A_j(\oi,:) \bigg],
\end{align}
where $\y_i^{\backslash j} = \y_i - \Ahi \mhi - \sum_{j' \neq j} w_{ij'} \A_{j'}(\oi,:) \m_{j'}$.

%-------------------------------------------
%  derivatives of q(v_i)
\newcommand{\Lhi}{\calL^h_i}
\noindent As a function of the parameters of $q(\v_i)$, the lower bound $\calL$ is:
\begin{align}
\nonumber
\Lhi \define&
 \log \Normal(\y_i; \sum_{j=1}^Q w_{ij} \A_j(\oi,:) \m_j + \Ahi \mhi, \beta_i^{-1} \I)
 - \frac{1}{2} \beta_i \trace \tilde{\K}^h_i(\oi,\oi)
 - \frac{1}{2} \beta_i \trace \Shi (\Ahi)^T \Ahi
 \\
  &- \frac{1}{2} \log |k(\Zhi,\Zhi) (\Shi)^{-1}| -\frac{1}{2} \trace k(\Zhi,\Zhi)^{-1} (\mhi (\mhi)^T + \Shi) ,
\end{align}

\noindent The derivatives of $\Lhi$ wrt $\mhi$ and $\Shi$ are given by:
\newcommand{\ynoh}{\y_i^{\backslash h}}
\begin{align}
\deriv{\Lhi}{\mhi}
& = \beta_i \A_i^T \ynoh - \left[k(\Zhi,\Zhi)^{-1} +  \beta_i \A_i^T \A_i \right] \m_i \\
\deriv{\Lhi}{\Shi} 
&= \frac{1}{2} \S_i^{-1} - \frac{1}{2} \left[ k(\Zhi,\Zhi)^{-1} + \beta_i \A_i^T \A_i \right] ,
\end{align}
where $\ynoh = \y_i - \sum_{j=1}^Q w_{ij} \A_j(\oi,:) \m_j$.

% comment on computation
\noindent It can be seen that the derivatives of the parameters of $q(\v_i)$ only involve the observations of the output dimension $i$.
The derivatives of the parameters of $q(\u_j)$ involve the observations across all output dimensions but decompose as a sum of contributions from individual outputs.
Therefore, computation of the derivatives (and hence the update equations) can be distributed or parallelized easily.
This attractive property allows the model to scale to a very large number of inputs and outputs.

% Comment on other hyperparameters

%\subsubsection{Update Equations}
% and comment on the intuition of the update equations: e.g. y(x) - g(x) for h(x) 
\subsection{Prediction}
The predictive distribution of the $i$-th output for a test input $\x_*$ is 
\begin{align}
p(\fstar | \y, \x_*) = \int \Normal(\fstar; \sum_{j=1}^Q w_j g_{j*} + h_{i*}, 0) p(\g_* | \y, \x_*) p(h_{i*} | \y, \x_*) \der \g_* \der h_{i*},
\end{align}
where $p(\g_* | \y, \x_*) = \prod_{j=1}^Q \Normal(g_{j*}; \mu_{j*}, s_{j*})$ and $p(h_{i*} | \y, \x_*) = p(h_{i*}; \mu^h_{i*}, s^h_{i*})$ are the predictive distributions of the sparse GPs as given in eq. \ref{eq:sorprediction}.
Therefore we have:
\begin{align}
p(\fstar | \y, \x_*) = \Normal(\fstar; \sum_{j=1}^Q w_{ij} \mu_{j*} + \mu_{i*}, w_{ij}^2 s_{j*} + s_{i*}). 
\end{align}

%\begin{linenomath}
%\begin{align}
%\sum_{i,h} \lambda_i \lambda_h cov [f(\x, \x')]
%&= \sum_{i,h}   \sum_{j,j'} \lambda_i g(\x_i - \vs_j) k(\vs_j, \vs_{j'}) \lambda_h g(\x_h - \vs_j') \\
%\end{align}
%\end{linenomath}

\section{Toy Experiments}
The first toy experiment uses two identical outputs which are noisy version of the same version plus some noise: $y_1(x) = sin(x) + \epsilon$ and $y_2(x) = sin(x) + \epsilon$, $\epsilon \sim \Normal(0,0.01)$.
In this case the shared function $g(x)$ is $sin(x)$ and the independent processes $h_1(x)$ and $h_2(x)$ are just white noises.
Each output has missing values in one region of the input space.
The predictive distributions of the multiple-gp model compared to that of independent gps are shown in \ref{fig4}.

\noindent The second toy experiment uses similar setting as the first one, except that now $y_1(x) = sin(x) + \epsilon$ and $y_2(x) = -sin(x) + \epsilon$. 
In this case the shared function is still $g(x) = sin(x)$, but the weights should be opposite i.e. $w_1 = 1$ and $w_2 = 1$.
The learning procedure was indeed able to recovered this relation and learned that $w_1 = 1.2$ and $w_2 = -1.3$.
The predictive distributions are shown in \ref{fig5}.

\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{figures/ssvi-y1.eps} &
\includegraphics[scale=0.5]{figures/ssvi-svi1.eps} \\
\includegraphics[scale=0.5]{figures/ssvi-y2.eps} &
\includegraphics[scale=0.5]{figures/ssvi-svi2.eps} \\
\multicolumn{2}{c}{\includegraphics[scale=0.5]{figures/ssvi-y1byg.eps} }
\end{tabular}
\label{fig4}
\caption{Predictive distributions of the multipe-output gps (left column) and independent gps (right column) for the first toy example. The predictive distribution by $g(x)$ for $y_1(x)$ is shown in the bottom figure.}
\end{figure*}

\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{figures/ssvi2-y1.eps} &
\includegraphics[scale=0.5]{figures/ssvi2-svi1.eps} \\
\includegraphics[scale=0.5]{figures/ssvi2-y2.eps} &
\includegraphics[scale=0.5]{figures/ssvi2-svi2.eps} \\
\multicolumn{2}{c}{\includegraphics[scale=0.5]{figures/ssvi-y1byg.eps} }
\end{tabular}
\label{fig5}
\caption{Predictive distributions of the multipe-output gps (left column) and independent gps (right column) for the second toy experiment. The predictive distribution by $g(x)$ for $y_1(x)$ is shown in the bottom figure.}
\end{figure*}

\section{Appendix 1: Derivatives}
Here we derive the gradients of the lower bound wrt the  hyperparameters.
As a template, consider the lower bound in Hensman et al (as a function of the hyperparameters):
\begin{align}
\nonumber
\calL
=& \log \Normal(\y; \K_{NM} \K_{MM}^{-1} \m, \beta^{-1}\I)
 - \frac{1}{2} \beta \trace \tilde{\K}
 - \frac{1}{2} \beta \trace (\S\K_{MM}^{-1} \K_{MN} \K_{NM} \K_{MM}^{-1}) \\  \nonumber
&- \frac{1}{2} \left( \log |\K_{MM}| + \trace(\K_{MM}^{-1}(\m \m^T + \S)) \right) \\ \nonumber
=& \underbrace{\log \Normal(\y; \A \m, \beta^{-1}\I)}_{\calL_1}
 - \underbrace{\frac{1}{2} \beta \trace (\K_{NN} - \A\K_{MN})}_{\calL_2}
 - \underbrace{\frac{1}{2} \beta \trace (\S\A^T\A)}_{\calL_3} \\  
&- \underbrace{\frac{1}{2} \left( \log |\K_{MM}| + \trace(\K_{MM}^{-1}(\m \m^T + \S)) \right)}_{\calL_4},
\end{align}
where $\A = \K_{NM} \K_{MM}^{-1}$.
Notice that we have re-written the sum of individual terms in matrix form which will make the derivation and also computation easier.

\subsection{Derivative of the Noise Hyperparameter}
The derivative of the noise hyperparameter $\beta$ is easily computed as:
\begin{align}
\deriv{\calL}{\beta} = \frac{N}{2\beta} - \frac{1}{2} (\y - \A\m)^T (\y - \A\m) - \frac{\calL_2}{\beta} - \frac{\calL_3}{\beta}.
\end{align}
\subsection{Derivatives of the Covariance Hyperparameters}  
To simplify the math, we utilize the matrix $\A$ defined above.
Firstly, the derivative of $\A$ wrt a covariance hyperparameter $t$ is given by:
\begin{align}
\deriv{\A}{t} = \left(\deriv{\K_{NM}}{t} - \A \deriv{\K_{MM}}{t}\right)\K_{MM}^{-1}.
\end{align}
The derivatives of $\calL_1, \calL_2, \calL_3 \text{ and } \calL_4$ are thus given by:
\begin{align}
\deriv{\calL_1}{t} &= \beta (\y - \A\m)^T \deriv{\A}{t} \m \\
\deriv{\calL_2}{t} &= \frac{1}{2}\beta \trace \left(\deriv{\K_{NN}}{t} - \A \deriv{\K_{MN}}{t} - \deriv{\A}{t} \K_{MN}\right) \\
\deriv{\calL_3}{t} &= \beta \trace \left(\A \S \deriv{\A^T}{t} \right) \\
\deriv{\calL_4}{t} &= \frac{1}{2}  \trace \left(\K_{MM}^{-1} \deriv{ \K_{MM}}{t}\right) - \frac{1}{2} \trace \left(\K_{MM}^{-1} \deriv{\K_{MM}}{t} \K_{MM}^{-1} (\m \m^T + \S) \right) 
\end{align}
The derivatives are then computed by taking the derivatives of the covariance matrices $\K_{NN} (\text{the diagonal only}), \K_{NM} \text{ and }  \K_{MM}$, hence the covariance function, wrt the hyperparameters.

\subsection{Derivatives of the Inducing Inputs}
To compute the derivatives of $\calL$ wrt the inducing inputs, first notice that $\Z = \{\z_m\}_{m=1}^M$ are also parameters of the covariance matrices $\K_{NM}$ and $\K_{MM}$.
Hence the derivative wrt a single dimension of an inducing input, i.e. $z_{mj}$, is the same as that of $\deriv{ \calL}{t}$.

%Since $MD$ parameters are needed for the inducing inputs, it appears that the derivatives of all inducing inputs would require $\calO(MD \times M^3)$ in computation.
%However, this complexity can actually be reduced to $\calO(DM^3)$ using the following lemma. \\
%
%\noindent \textbf{Lemma} Let $A, B$ be two matrices of size $N \times M$ and $M \times N$, respectively. Furthermore, $B$ has the property that only one of its rows or columns is non-zero. The complexity of $\trace(AB)$ is only $\calO(N)$. \\
%
%\noindent \textbf{Proof} Let $m <= M$ be the non-zero row of $B$. We have
%\begin{align}
%\trace (AB) = \sum_{i=1}^N \sum_{j=1}^M A_{ij} B_{ji} = \sum_{i=1}^N A_{im} B_{mi},
%\end{align}
%which clearly takes $\calO(N)$. It is easy to see that the lemma also holds when $B$ is symmetric and only one of its row and the corresponding column is non-zero. \\
%
%\noindent To exploit the property in the Lemma, we re-write $\frac{d \calL_1}{dt}, \frac{d \calL_2}{dt}, \frac{d \calL_3}{dt}, \frac{d \calL_4}{dt}$ by expanding $\frac{d\A}{dt}$ (here $t = z_{mj}$):

\noindent We re-write $\deriv{\calL_1}{t}, \deriv{\calL_2}{t}, \deriv{ \calL_3}{t}, \deriv{\calL_4}{t}$ by expanding $\deriv{\A}{t}$ (here $t = z_{mj}$):
\begin{align}
\nonumber
% dL1
\deriv{\calL_1}{t}
 &= \beta \trace (\y - \A\m)^T \left(\deriv{\K_{NM}}{t} -  \A \deriv{ \K_{MM}}{t}\right)\K_{MM}^{-1} \m \\
&= \beta \trace \K_{MM}^{-1} \m (\y - \A\m)^T \deriv{\K_{NM}}{t} 
-\beta \trace \K_{MM}^{-1} \m (\y - \A\m)^T \A \deriv{\K_{MM}}{t} \\
%dL2
\deriv{\calL_2}{t}
&= - \beta \trace \A^T \deriv{\K_{NM}}{t}
 + \frac{1}{2} \beta \trace \A^T \A \deriv{\K_{MM}}{t}  \\
% dL3
\deriv{\calL_3}{t}
&= \beta \trace \K_{MM}^{-1} \S \A^T \deriv{\K_{NM}}{t}
 - \beta \trace \K_{MM}^{-1} \S \A^T \A \deriv{\K_{MM}}{t}\\
% dL4
\deriv{\calL_4}{t}
 &= \frac{1}{2}  \trace \K_{MM}^{-1} \deriv{\K_{MM}}{t}
  - \frac{1}{2} \trace \K_{MM}^{-1} (\m \m^T + \S) \K_{MM}^{-1} \deriv{\K_{MM}}{t} 
\end{align}
From the above equations we get,
\begin{align}
\deriv{\calL}{t} = \trace \D_1 \deriv{\K_{NM}}{t} + \trace \D_2 \deriv{ \K_{MM}}{t},
\end{align}
where 
\begin{align}
\D_1 =& \beta \K_{MM}^{-1} \m (\y - \A\m)^T
 + \beta \A^T
 - \beta \K_{MM}^{-1} \S \A^T \\ \nonumber
\D_2 =& -\beta \trace \K_{MM}^{-1} \m (\y - \A\m)^T \A
 - \frac{1}{2} \beta \A^T \A
  + \beta \K_{MM}^{-1} \S \A^T \A	 \\ 
  &-\frac{1}{2} \K_{MM}^{-1} + \frac{1}{2} \K_{MM}^{-1} (\m \m^T + \S) \K_{MM}^{-1}
\end{align}
Notice that $\D_1$ and $\D_2$ can be pre-computed with a cost of $\calO(M^3)$ (or $\calO(N_bM^2)$ if the minibatch size $N_b > M$).
The computational cost of taking derivatives of $MD$ inducing parameters is thus $\calO(M^3 + MDM) = \calO(M^3)$ as the cost of the two $\trace$ operators is $\calO(M)$ due to the fact that only $\calO(M)$ elements of $\deriv{\K_{MM}}{t}$ or $\deriv{\K_{NM}}{t}$ are non-zero.
This fact can be further exploited to perform vectorized operation, for e.g. in Matlab, such that the iteration over all inducing inputs can be avoided.

\section{Appendix 2: Gaussian Identity}
Let $\y$, $\g = \{\g_j\}_{j=1}^q$, and $\h$ be random variables with multivariate Gaussian distributions: 
$p(\y | \g, \h) = \Normal(\y; \sum_{j=1}^Q \W_j \g_j + \W \h, \beta^{-1} \I)$, $p(\g_j) = \Normal(\g_j; \m_j, \S_j)$, and $p(\h) = \Normal(\h; \m, \S)$.
The following identity is important in deriving the evidence lower bound:
\begin{align}
\nonumber
\int \log p(\y | \g, \h) \prod_{j=1}^q p(\g_j) p(\h) \der \g \der \h
=& \log \Normal(\y; \sum_{j=1}^q \W_j \m_j + \W \m, \beta^{-1} \I) - \frac{1}{2} \beta \trace \W^T \W \S \\
\label{eq:identity}
&- \frac{1}{2} \beta \trace \sum_{j=1}^q \W_j^T \W_j \S_j.
\end{align}
The identity can be proved by using this fact: 
\begin{align}
\nonumber
\int (\W\x - \Mu)^T \BigSigma^{-1} &(\W \x - \Mu) \Normal(\x; \m, \S) \der \x \\
\nonumber
&= (\Mu - \W\m)^T \BigSigma^{-1} (\Mu - \W\m) + \trace \W^T \BigSigma^{-1} \W \S.
\end{align}

\section{Appendix: Analysis on Learning of the Inducing Inputs}
In this section we present some analysis on learning of the inducing inputs under the stochastic variational inference procedure on some toy problems.
We use three real-valued truth functions of scalar inputs: the first one was used in Snelson et al, the second one is a function $f(x) = \sin(x) + \cos(x)$, and the third one is a sample function generated from a GP with squared exponential with ARD covariance function with a lengthscale of 2 and signal variance of 1.

\noindent \paragraph{Experimental Settings} We use gpsvi to optimize for \textit{all} of the parameters in the model, i.e. the variational parameters of the posterior, the covariance hyperparameters, the noise hyperparameter, and the inducing inputs.
For the hyperparameters, we use a momentum of $0.9$ and a fixed learn rate of $1e-5$. 
For the variational parameters, we use a learn rate of $0.01$ (no momentum was used).
We vary the learn rate of the inducing inputs from $1e{-2}$ to $1e{-6}$.
Using momentum for the inducing inputs does not seem to have much effects.
The maximum number of iterations is 500, the batch size 5, and the number of inducing values 10.

% effect of learn rate on inducing inputs
\noindent \paragraph{Effects of the Learn Rates }
In Figure \ref{fig1}, \ref{fig2}, and \ref{fig3}, we show the lower bounds, predictive distributions, and the learned inducing inputs using stochastic variational inference with varying learn rates for the inducing inputs.
We also show the results with FITC for comparison.
It can be seen from the figures that, as typical in stochastic optimization, the behavior of GPSVI is very sensitive to the learn rate.
In particular, when the rate is slow e.g. $1e{-5}$, almost no progress was made in learning the inducing inputs.
However, when the rate increases to e.g. $1e{-3}$, the inducing inputs are spread out more evenly in the input space compared to the initial locations.
Although not shown the in figures, when the learn rate is too large, e.g. $1e{-02}$, the inducing inputs may spread to regions far outside of the training intervals.
An example of this phenomenon can be observed in the top left figure in Figure \ref{fig1}.
The learn rates also affect the evidence lower bound which seems to wiggle more compared to not learning the inducing inputs.
The predictive distributions are sensible particularly because the toy functions exhibit sparsity and are easy to learn.  \\

\noindent The results of this qualitative analysis do not seem to suggest that learning of the inducing inputs does not work in GPSVI.
Its effectiveness perhaps rests upon empirical performance on real datasets.

\noindent \paragraph{The Noise Parameter} It seems harder to set the initial value and the learning rate of the noise parameter compared to standard GP. 
While a large range of values of noise can be used in standard gp or fitc (e.g. 0.5), a very small value of noise $1e{-02}$, which is the true noise, is required for gpsvi to work well.
It appears that an adaptive learning rate may be required for the noise parameter.

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-03.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-04.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-05.eps} \\
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-03-bound.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-04-bound.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-05-bound.eps} \\ 
(a) learn rate = $1e{-3}$ & (b) learn rate = $1e{-4}$ & (c) learn rate = $1e{-5}$ \\
\multicolumn{3}{c}{\includegraphics[scale=0.4]{figures/func1-fitc.eps}}
\end{tabular}
\caption{Predictive distributions and learned inducing inputs by GPSVI and FITC for the first function. Top row: GPSVI with different learning rates (the rates decreases from left to right). Bottom row: FITC. Magenta dots : training points. Solid blue line: predictive mean. Grey-shaded area and solid black lines: two standard deviations of the predictive distributions. Black (+) crosses: initial locations of inducing inputs. Magenta (+) crosses: learned locations of inducing inputs.  Middle row: the evidence lower bound of gpsvi vs. iteration.}
\label{fig1}
\end{figure*}

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-03.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-04.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-05.eps} \\
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-03-bound.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-04-bound.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-05-bound.eps} \\ 
(a) learn rate = $1e{-3}$ & (b) learn rate = $1e{-4}$ & (c) learn rate = $1e{-5}$ \\
\multicolumn{3}{c}{\includegraphics[scale=0.4]{figures/func2-fitc.eps}}
\end{tabular}
\caption{Predictive distributions and learned inducing inputs by GPSVI and FITC for the second function. The legends are same as in Figure 1.}
\label{fig2}
\end{figure*}

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-03.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-04.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-05.eps} \\
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-03-bound.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-04-bound.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-05-bound.eps} \\ 
(a) learn rate = $1e{-3}$ & (b) learn rate = $1e{-4}$ & (c) learn rate = $1e{-5}$ \\
\multicolumn{3}{c}{\includegraphics[scale=0.4]{figures/func3-fitc.eps}}
\end{tabular}
\caption{Predictive distributions and learned inducing inputs by GPSVI and FITC for the third function. The legends are same as in Figure 1.}
\label{fig3}
\end{figure*}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}