### 2014-03-18 paper/model.tex

Before diving into details of the specification, we discuss the modeling philosophy behind our collaborative multioutput Gaussian processes.
To learn the outputs jointly, we need a mechanism through which information can be transferred among the outputs.
This is achieved in the model by allowing the outputs to share multiple sets of inducing variables, each of which captures a different pattern common to the outputs.
%For exposition, consider a set of inducing points for a moment.
%Because they serve as the sufficient statistics of the individual outputs, a pattern common to the outputs can be encapsulated in these points when learning is done jointly. 
%Different patterns in the data can be captured by having  multiple set of inducing variables.
These variables indeed play a double pivotal role in the model: they collaborate sharing of information across the outputs and they provide the notion of  sufficient statistics that induces sparse processes.

\newcommand{\Zj}{\Z_j}
\newcommand{\Zhi}{\Z^h_i}
Consider multiple output regression of $P$ tasks with inputs $\X = \{\x_n \in \calR^D\}_{n=1}^N$ and outputs $\y = \{\y_i\}_{i=1}^P$ where $\y_i = \{y_{in}\}_{n=1}^N$.
We shall model each output as a weighted combination of $Q$ shared latent functions, plus an individual latent function unique to that output for greater flexibility.
The $Q$ shared functions have independent Gaussian process priors, $g_j(\x) \sim \GP(0, k_j(\cdot,\cdot))$, $j= 1 \hdots Q$.
As we wish to use sparse processes, we introduce the set of \emph{inducing variables} $\u_j$ for  each $g_j(\x)$, i.e., $\u_j$ contains the values of $g_j(\x)$ at the inducing inputs $\Z_j$.
Similarly, each individual function of an output also has GP prior, i.e., $h_i(\x) \sim \GP(0, k^h_i(\cdot,\cdot))$ with inducing inputs $\Zhi$ and inducing values $\v_i$.
The inducing inputs live in the same space as the inputs $\X$, which should be apparent since their corresponding values come from the processes defined on the input space.
For convenience, we shall assume all of the processes have the same number of inducing points, $M$, however we emphasize that this is \emph{not} imposed in practice.

% information is transfered via the inducing variables

As notations we denote the collective variables: $\g = \{\g_j\}$, $\h = \{\h_i \}$, $\u = \{\u_j\}$, $\v = \{\v_i\}$, $\Z = \{\Zj\}$, and $\Z^h = \{\Zhi \}$ where $\g_j = \{g_j(\x_n)\}$, $\h_i = \{h_i(\x_n)\}$. 
Note that we reserve subscript $i$ for indexing the outputs and their corresponding individual processes ($i = 1 \hdots P$), $j$ for the latent processes ($j = 1 \hdots Q$), and $n$ for the inputs ($n = 1 \hdots N$).

From the definition of the GPs and the independence of the processes, the \emph{prior} of the multioutput model can be written as:
\begin{align}
\label{eq:gu}
p(\g | \u) &= \prod_{j=1}^Q p(\g_j | \u_j) = \prod_{j=1}^Q \Normal(\g_j; \BigMu_j, \tilde{\K}_j )\\
\label{eq:u}
p(\u) &= \prod_{j=1}^Q p(\u_j) = \prod_{j=1}^Q \Normal(\u_j; \vec{0}, k(\Zj, \Zj)) \\
\label{eq:hv}
p(\h | \v) &= \prod_{i=1}^P p(\h_i | \v_i) = \prod_{i=1}^P \Normal(\h_i; \BigMu^h_i, \tilde{\K}^h_i)\\
\label{eq:v}
p(\v) &= \prod_{i=1}^P p(\v_i) = \prod_{i=1}^P \Normal(\v_i; \vec{0}, k(\Zhi, \Zhi)),
\end{align}
where
\begin{align}
 \BigMu_j &= k(\X,\Zj)k(\Zj,\Zj)^{-1}\u_j \\
\BigMu^h_i &= k(\X,\Zhi)k(\Zhi,\Zhi)^{-1}\v_i \\
\tilde{\K}_j &= k_j(\X,\X) - k(\X,\Zj)k(\Zj,\Zj)^{-1}k(\Zj,\X) \\
\tilde{\K}^h_i &= k^h_i(\X,\X) - k(\X,\Zhi)k(\Zhi,\Zhi)^{-1}k(\Zhi,\X).
\end{align}
In the equations and hereafter, we omit the subscripts $j,h,i$ from the kernels $k_j(\cdot,\cdot)$ and $k^h_i(\cdot,\cdot)$ when it is clear from the parameters inside the parentheses which covariance function is in action.
Here, equations \ref{eq:u} and \ref{eq:v} follow directly from the properties of GPs, while the expressions for $p(\g|\u)$ and $p(\h|\v)$ (eq. \ref{eq:gu} and \ref{eq:hv}) come from the conditionals of the multivariate Gaussian distributions.
Instead of writing the joint priors $p(\g,\u)$ and $p(\h,\v)$, the above equivalent equations are given to emphasize the important role of the inducing variables in the model.

The \emph{likelihood} with standard iid Gaussian likelihood is given by:
\begin{align}
p(\y | \g, \h ) = \prod_{i=1}^P \prod_{n=1}^N \Normal( y_{in} ; \sum_{j=1}^Q w_{ij} g_j(\x_n) + h_i(\x_n), \beta_i^{-1}),
\end{align}
which says that $y_{in}$ is a linear combination of the latent functions with weight $w_{ij}$ for the $j$th shared latent process plus a contribution from the individual process $h_i$.
As the latent values $\g$ are specified conditioned on the inducing variables $\u$, this construction implies that each output is a weighted combination of the inducing values.
We note that if $\u$ and $\v$ are marginalized out, we obtain the semiparametric latent factor models \citep{teh-et-al-aistats-05}.
However, doing so is against the purpose of this model which encourage sharing of outputs via the inducing variables.
Furthermore, as we shall see in the next section, explicit representation of these variables is instrumental to scalable inference of the model.

%\textbf{Augmented sparse GPs}
%Toward scalable modeling, we replace standard GPs with sparse GPs augmented with \textit{different} set of inducing inputs.
%This adds much flexibility to the model as each of the shared process $g_j(\x)$ can model a different pattern in the data with its own covariance function and inducing inputs. The roles of $g_j(\x)$ and $h_i(\x)$ can be quite different, so it is necessary that each has its own inducing inputs.
%Furthermore, $g_j(\x)$ can be seen as a \textit{global} function operating on the entire input space (of all output dimensions), while each $h_i(\x)$ operates only on the inputs of the $i$-th output, which can be a subspace of the input.