### 2014-03-19 paper/paper.tex

% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two
% column, 8 1/2 by 11 inch proceedings format.

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage{latexsym,amsbsy,amssymb,color,xspace,booktabs}

\include{macros}
\include{local_macros}

\title{Collaborative Multi-output Gaussian Processes}

\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\section{INTRODUCTION}
\input{intro}

\section{MODEL SPECIFICATION \label{sec:model}}
\input{model}

\section{INFERENCE \label{sec:inference}}
\input{inference}

\section{EXPERIMENTS \label{sec:experiments}}
\input{experiments}

\section{DISCUSSION \label{sec:discussion}}
\input{discussion}

%\subsubsection*{Acknowledgements}

\bibliographystyle{apalike}
\bibliography{references}

\end{document}